`flippant` -- An `R` package for the automated analysis of dithionite scramblase

### 2015-09-23 inst/manuscript/manuscript.rmd
Figure reformatting to a 3/5-2/5 split;
Further editing;

figure1 <- multipanelfigure(
  width = 17.22, 
  height = 18, 
  units = "cm", 
  columns = 2, 
  rows = 2,
  interPanelSpacing = 0.25)

@@

Numeric scramblase assay results such as $\tau$ may be generated in tabular form
using
```{r, code = formatR::usage('scramblaseAssayStats'), eval = FALSE}
```
with the same modifying arguments to the function as for `scramblaseAssayPlot`.

@@

`flippant`s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
step plots the corresponding spectral traces (Figure 1A). Note that alorithmic
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments.
```{r, eval = FALSE}
# Plot spectral traces (Figure 1A)
scramblaseAssayTraces("inputTable.txt", timeMax = 350)
```
```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.cap = "**`flippant`-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1, 
  leftPanel = 1, 
  rightPanel = 2, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMax = 350)
)
grid.draw(figure1)
```

#Conclusion
```{r, echo = FALSE, cache = TRUE}
setwd(tmpDir)
analysis_timing_n <- 100
analysis_timing_function <- function(){

  scramblaseAssayTraces("inputTable.txt", timeMax = 350)
}
analysis_timing <- microbenchmark(
  analysis_timing_function(),
  times = analysis_timing_n)

@@

analysis_timing_iqr <- round(
  diff(
    unlist(
      summary(analysis_timing)[c("lq", "uq")],
      use.names = FALSE)),
  2)
```
With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
`R` [@r_core_team_r:_2015] as an environment for statistical computing and
graphics, scientists can quickly analyse such data and arrive at 
publication-grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a system running `Linux`
and `R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr` \text{s}$; $n=`r analysis_timing_n`$).

>>>>

figure1 <- multipanelfigure(
  widths = rep(17.22/3-2*0.25, 3), 
  heights = rep(18/5-4*0.25, 5), 
  units = "cm",
  interPanelSpacing = 0.25)

@@

Numeric scramblase assay results such as $\tau$/$\alpha$ may be generated in 
tabular form using
```{r, code = formatR::usage('scramblaseAssayStats'), eval = FALSE}
```
with the same modifying arguments to the function as for `scramblaseAssayPlot`.

@@

`flippant`s capabilities are demonstrated here by reanalyzing data underlying 
Figure 4C/D in @ploier_perturbed_submitted. The following work flow in a first 
step plots the corresponding spectral traces (Figure 1A). Note that alorithmic
alignment to the time of dithionite addition is employed (default) as well as 
custom trimming of the time axis to account for divergent acquisition 
times between experiments.
```{r, eval = FALSE}
# Plot spectral traces (Figure 1A)
scramblaseAssayTraces("inputTable.txt", timeMax = 350)
```
```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "**`flippant`-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`, while (B) shows protein to phospholipid ration (PPR) plots for the data using a x-axis trimmed to a maximum value of 2 (see text for details)."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1,
  bottomPanel = 3,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMax = 350) + theme_bw())
figure1 <- addpanel(
  figure = figure1,
  topPanel = 4,
  bottomPanel = 5,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayPlot("inputTable.txt") + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
grid.draw(figure1)
```

Scramblase activity analysis as described above is performed on the data set with
the following calls.
```{r, eval = FALSE}
# Plot PPR plot (Figure 1B)
scramblaseAssayPlot("inputTable.txt") + xlim(0, 2)
# Write out results in tabular manner
scramblaseAssayStats("inputTable.txt")
```
```{r, echo = FALSE}
setwd(tmpDir)
max_PPR <- read.delim(
    "inputTable.txt",
    check.names = FALSE) %>%
  flippant:::calculatePpr() %>%
  max(
    na.rm = TRUE) %>%
  round(
    digits = 2
  )
```
The first call results in the protein to phospholipid ratio (PPR) plot shown in
Figure 1B. Using the customization facilities provided by `ggplot2` 
[@wickham_ggplot2_2009] as a plotting system the plot's x-axis is trimmed to
avoid visual compression of all data by a single data point from the wild type 
rhodopsin series in experiment 1, where $\text{PPR}=`r max_PPR`$. Dashed lines
are added automatically as graphical representations of the fit constant 
$\alpha$ or $\tau$, but are difficult to evaluate visually.

Recapitulation of the calculations underlying PPR plot generation by 
`scramblaseAssayPlot` in a call to `scramblaseAssayStats` produces tabular 
output represented in Table 1, which includes series-specific fit constants 
$\alpha$ in numeric and easily accessible form.
```{r, echo = FALSE}
setwd(tmpDir)
ppr_results <- scramblaseAssayStats("inputTable.txt")
ppr_results_tau_average <- ppr_results %>%
  split(ppr_results$`Experimental Series`) %>% 
  sapply(function(y){mean(y$`Fit Constant`, na.rm = TRUE)}) %>%
  round(2)
ppr_results_tau_average_names <- names(ppr_results_tau_average) %>%
  head(n = length(ppr_results_tau_average) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(names(ppr_results_tau_average), n = 1))
ppr_results_tau_average_print <- ppr_results_tau_average %>%
  unname() %>%
  head(n = length(ppr_results_tau_average) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results_tau_average, n = 1))
knitr::kable(
  ppr_results,
  row.names = FALSE,
  caption = "**Tabular output of `flippant`-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** An `ASCII`-version of this output is produced by calling `scramblaseAssayStats`.")
```

Mean $\tau$ values (or fit constants $\alpha$) correspond with `r ppr_results_tau_average_print` for `r ppr_results_tau_average_names`, 
respectively, very well to the prior manual analyis by 
@ploier_perturbed_submitted, where $\tau \approx 0.12$ was reported for mustant 
and $\tau \approx 0.26$ for wild type rhodopsin. Based on the manual analysis of
the data we conclude in @ploier_perturbed_submitted that the mutant rhodopsins
investigated are reconsituted into proteoliposomes in monomeric form, contrary 
to the wild type, which is reconstituting as a homodimer. `flippant` analysis of
the data, shows identical trends and close matches of the results, supporting 
identical conclusions.

#Conclusion
```{r, echo = FALSE, cache = TRUE}
setwd(tmpDir)
analysis_timing_n <- 100
analysis_timing_function <- function(){
  scramblaseAssayTraces(
    "inputTable.txt",
    timeMax = 350)
  scramblaseAssayPlot("inputTable.txt")
  scramblaseAssayStats("inputTable.txt")
}
analysis_timing <- microbenchmark(
  analysis_timing_function(),
  times = analysis_timing_n)

@@

analysis_timing_iqr <- round(
  diff(
    unlist(
      summary(analysis_timing)[c("lq", "uq")],
      use.names = FALSE)),
  2)
```
With `flippant` we present an integrated solution for data analysis in the 
context of dithionite scramblase assays. Requiring only basic familiarity with 
`R` [@r_core_team_r:_2015] as an environment for statistical computing and
graphics, scientists can quickly analyse such data and arrive at 
publication-grade graphics that offer extensive facilities for individual 
optimization and adaptation [@wickham_ggplot2_2009]. On a laptop system with 
`r parallel::detectCores()` cores and 8 GB RAM (no SSD), running `Linux` and 
`R` version `r paste(R.version[c("major","minor")], collapse = ".")`, the
statements generating equivalents to the plots and calculations presented here
take a mere `r analysis_timing_median` seconds (median; 
$\text{IQR}=`r analysis_timing_iqr`~\text{s}$; $n=`r analysis_timing_n`$).
